Making ice sheet models \\ scale properly

### 2023-01-06 talk/slides.tex
progress

\medskip
\item<2> {\footnotesize see UAF's \href{https://pism.io/}{Parallel Ice Sheet Model (\texttt{pism.io})} for a model which does \emph{not} make these simplifications}
\item<2> {\footnotesize next, a simple definition of ``ice sheet model''}
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-glacier-ice.png}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{notation for the simplified model}

@@

>>>>

\medskip
\item<2> {\footnotesize see UAF's \href{https://pism.io/}{Parallel Ice Sheet Model (\texttt{pism.io})} for a model which does \emph{not} make these simplifications}
\item<2> {\footnotesize next, a basic definition of ``ice sheet model''}
\end{itemize}
\end{column}
\begin{column}{0.42\textwidth}
\hfill \includegraphics[width=0.9\textwidth]{images/map-glacier-ice.png}
\end{column}
\end{columns}
\end{frame}

@@

\begin{frame}{notation for the basic model}

@@

\begin{frame}{shallow ice sheet models}

\begin{itemize}
\item the model on the last slide is \emph{not yet} in common use for ice sheet simulations, because for now it is expensive to compute
    \begin{itemize}
    \item goal: make it fast enough to use!
    \end{itemize}
\item current-technology models use various shallow simplifications
\item simplest is a lubrication approximation, the \alert{shallow ice approximation}, in which $- \bu|_s \cdot \bn_s$ resolves to a nonlinear differential operator  
\end{itemize}
\end{frame}

\begin{frame}{shallow ice approximation: details}

{\small
\begin{align*}
s - b &\ge 0 \\
\frac{\partial s}{\partial t} + \Phi(s) - a &\ge 0 \\
(s - b) \left(\frac{\partial s}{\partial t} + \Phi(s) - a\right) &= 0
\end{align*}
where
\begin{align*}
\Phi(s) &= - \frac{\gamma}{\pp} (s-b)^{\pp} |\grad s|^{\pp} \\
        &\qquad\qquad - \grad \cdot\left(\frac{\gamma}{\pp+1} (s-b)^{\pp+1} |\grad s|^{\pp-2} \grad s\right)
\end{align*}
here $\nabla$ only acts in horizontal variables, and constants $\pp = \nn+1$ and $\gamma > 0$ are related to ice deformation

\begin{itemize}
\item $\Phi$ is a \alert{doubly-nonlinear differential operator}, simultaneously of porous medium and $\pp$-Laplacian type
\end{itemize}
}
\end{frame}
